---
title       : Predicting Car Acceleration 
subtitle    : Random Forest Model of the Auto Data set from the ISLR package
author      : allan cuenca
job         : 
framework   : revealjs        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Predicting Car Acceleration

#### A Random Forest Prediction Exercise


for the Johns Hopkins Bloomberg School of Public Health
<p><span style="font-size:14pts;">Data Science Specialization -- Developing Data Products
<p>via Coursera.org</span>

May 2015

---

## What does it do?

This [Shiny App](https://kawenks.shinyapps.io/JHBSPH_DevDataProd) uses the [Auto](http://www.inside-r.org/packages/cran/ISLR/docs/Auto) dataset in the [ISLR](http://cran.r-project.org/web/packages/ISLR/ISLR.pdf) package. It sets up a Random Forest prediction model to determine a car's acceleration based on several attributes.  

---

## Data Exploration - Auto dataset

1. 392 observations (vs. 32--mtcars), 9 variables (vs. 11--mtcars). 
2. 2 numeric variables (year, origin) that should be ordinal
3. 3 variables have relatively higher correlation to acceleration
<div style="float:center;">
```{r echo=F,render=T,fig.height=4}
par(mfrow=c(1,3))
plot(x=Auto$acceleration,y=Auto$horsepower,main="Horsepower")
plot(x=Auto$acceleration,y=Auto$displacement,main="Displacement")
plot(x=Auto$acceleration,y=Auto$weight,main="Weight")
```
</div>


---

## The prediction model

Feature Selection and Cross Validation Fine-Tuning

* horsepower  
* displacement  
* weight  
* cylinders
<p>
<p>
<div style="height:18pts">&nbsp;</div>
Fine-tune with 10-fold cross validation</p>



---

## Sample Error Rate


Random Forest has lower RMSE and explains more of the data variability.

```{r echo=FALSE, message=FALSE, warning=FALSE, render=FALSE,cache=TRUE}
library(randomForest)
library(dplyr)
library(ISLR)
library(caret)

data(Auto)

d <- Auto
d$cylinders <- as.factor(d$cylinders)

trainctrl <- trainControl(method="cv", number=10)
trix <- createDataPartition(d$acceleration, p=0.9,list=F)
set.ts <- d[-trix,]
set.tr <- d[trix,]

set.seed(12345)
fit.rf <- train(acceleration ~ horsepower + 
                    displacement + 
                    weight + 
                    cylinders,
                data=set.tr,
                method="rf",
                trControl=trainctrl,
                verbose=F)

fit.gbm <- train(acceleration ~ horsepower + 
                    displacement + 
                    weight + 
                    cylinders,
                data=set.tr,
                method="gbm",
                trControl=trainctrl,
                verbose=F)

fit.gam <- train(acceleration ~ horsepower + 
                    displacement + 
                    weight + 
                    cylinders,
                data=set.tr,
                method="gam",
                trControl=trainctrl,
                verbose=F)

r.rf <- fit.rf$results %>% filter(min(RMSE)==RMSE) %>% select(RMSE,Rsquared,RMSESD,RsquaredSD)
r.gbm <- fit.gbm$results %>% filter(min(RMSE)==RMSE) %>% select(RMSE,Rsquared,RMSESD,RsquaredSD)
r.gam <- fit.gam$results %>% filter(min(RMSE)==RMSE) %>% select(RMSE,Rsquared,RMSESD,RsquaredSD)

err.tbl <- rbind(r.rf,r.gbm,r.gam)
err.tbl <- cbind(Model=c("Random Forest","Bayesian Generalized Linear Model","Generalized Additive Model"),err.tbl)
```

```{r results='asis',echo=FALSE}
library(xtable)
names(err.tbl) <- c(" Model "," RMSE "," $R^2$ "," RMSE sd "," $R^2$ sd ")

err.disp <- xtable(err.tbl, digits=4,align=c(rep("left",2),rep("center",4)))
print(err.disp,type="html"
        , html.table.attribute='class=ErrRateTbl'
        ,include.rownames=F
        )

```






